---
title: "Reproducible Research Project 1"
author: "David R. Cote"
date: "6/23/2020"
output: html_document
---
The data for this assignment can be downloaded from the course web site:

Dataset: Activity monitoring data [52K]  
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as \color{red}{\verb|NA|}NA)  
date: The date on which the measurement was taken in YYYY-MM-DD format  
interval: Identifier for the 5-minute interval in which measurement was taken  
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

Loading Libraries:
```{r loadlibs, results="hide"}
library(data.table)
library(ggplot2)
library(knitr)
```

Loading Data:
```{r loaddata}
fileUrl <-  "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile=paste0(getwd(),'/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip")
```

Create Data Table:
```{r datatable}
ActivityData <- data.table::fread(input = "activity.csv")
```

Format Dates in Table:
```{r dateconv}
ActivityData$date<-as.Date(ActivityData$date)
```

Question 1: What is mean total number of steps taken per day?
